---
title: "simpleRegress"
author: "Alondra Cruz"
date: "Summer 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  
We are going to compute regression models finding one regression model that does the best in predicting the bear length, and its correspondence to the bears neck size, age, head width and head length. 
```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted r square that is the largest is the best. The best for the neck size is 0.7476.

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted r square for head length is 0.8434


```{r}
simp <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted r square for head width is 0.5321 

```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted r squared is 0.5073

```{r}
simp <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted r square 0.7869

```{r}
simp <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted r square 0.7423

### Conclusion
We had to compare all the models and order them worst to best based off their value of Adjusted-R Squared. Based of the list of Adjusted R- Squared values underneath, the best is head length. 

Age:           0.5073

Head Width:    0.5321

Weight:        0.7423

Neck:          0.7476

Chest:         0.7869

Head Length:   0.8434

 






















